---
title: "Text Analysis: The Alchemist by Paulo Coelho"
description: |
  Using text analysis to assess book sentiment.
author:
  - name: Laurel Abowd
date: 2021-03-13
output:
  distill::distill_article:
    self_contained: false
    code_folding: hide
---

```{r setup, include=FALSE, warning= FALSE, error = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(tidytext)
library(textdata)
library(pdftools)
library(ggwordcloud)
```

## **Introduction**

```{r, cache=TRUE}
# read in the data
alchemist <- pdf_text(here("data", "project-03", "the_alchemist_coelho.pdf"))

```

*The Alchemist* chronicles the fable of Santiago, a shepherd boy who travels through the desert to find treasure and fulfill his personal destiny. The journey takes unexpected turns along the way. The book is broken into three parts: Part One, Part Two, and the Epilogue. 

### **Citation:** 
Coelho, Paulo. The Alchemist. Translated by Alan R. Clarke. Published 1992. ISBN 0-7225-3293-8.
Retrieved from books-library.online here: 
https://books-library.online/files/books-library.online-12230111Ib9C3.pdf

```{r}
# Get text in data frame and tidy format

alchemist_tidy <- data.frame(alchemist) %>% 
  mutate(text_full = str_split(alchemist, pattern = '\\n')) %>% 
  unnest(text_full) %>% 
  mutate(text_full = str_trim(text_full)) 

```

## **Data Wrangling**

```{r}
# Remove introductory text and assign part 1, part 2, or epilogue to the text_full column
alchemist_df <- alchemist_tidy %>% 
  slice(-(1:7)) %>% 
  mutate(part = case_when(
    str_detect(text_full, "PART") ~ text_full,
    str_detect(text_full, "EPILOGUE") ~ text_full,
    TRUE ~ NA_character_
  )) %>% 
  fill(part) %>% 
  mutate(part = str_to_lower(part))
```

```{r message = FALSE}
# Get word counts by part

alchemist_tokens <- alchemist_df %>% 
  unnest_tokens(word, text_full) %>% 
  select(part, word)

alchemist_wordcount <- alchemist_tokens %>% 
  count(part, word)

# Remove stop words

alchemist_nonstop_words <- alchemist_tokens %>% 
  anti_join(stop_words) 

# "boy" appears frequently, but it is what the author uses to refer to the main character, Santiago.
# Include "boy" in the stop words

non_stop_counts <- alchemist_nonstop_words %>% 
  count(part, word) %>% 
  filter(word != "boy")

```

## **Word Count Analysis**

```{r}
# Find the top 5 words from each part

top_5_words <- non_stop_counts %>% 
  group_by(part) %>% 
  arrange(-n) %>% 
  slice(1:5)

# Reorder parts so that they occur in chronological order instead of alphabetical

top_5_words_chrn <- top_5_words %>% 
  mutate( part = factor(part,
         levels = c("part one", "part two", "epilogue"))) 


```

```{r fig.align= 'center'}
# Make graph
ggplot(data = top_5_words_chrn, aes(x = word, y =n)) +
  geom_col(fill = "gold") +
  facet_wrap(~part, scales = "free") +
  coord_flip() +
  labs(y = "word count",
       title = "The Alchemist top 5 words by part") +
  theme_grey()
```
**Figure 1.** Top five non-stop words in each part of *The Alchemist* 

These top five words reflect the story arc of *The Alchemist*. In Part One, Santiago is with his flock of sheep and dreams of treasure and a journey. In Part Two, Santiago travels through the desert and eventually meets the alchemist. Finally, in the epilogue, there is resolution of Santiago's personal destiny. 

## **NRC Sentiment Analysis** 

```{r message = FALSE}
# Sentiment analysis 

nrc <- get_sentiments(lexicon = "nrc")

alchemist_nrc <- alchemist_nonstop_words %>% 
  filter(word != "boy") %>%
  inner_join(nrc)

alchemist_nrc_counts <- alchemist_nrc %>% 
  count(part, sentiment)

# Get percentages of sentiments
alchemist_nrc_percent <- alchemist_nrc_counts %>% 
  group_by(part) %>% 
  mutate(percent = round((n/sum(n))*100, 2)) %>% 
  mutate( part = factor(part,
                        levels = c("part one", "part two", "epilogue"))) 
```

```{r fig.align = "center"}
# graph sentiments for the three parts

ggplot(data = alchemist_nrc_percent, aes(x = sentiment, y = percent)) +
  geom_col(fill = "gold") +
  facet_wrap(~part, scales = "free") +
  coord_flip() +
  labs(y = "percent occurrence",
       title = "The Alchemist NRC sentiment analysis by part") +
  theme_grey()


```
**Figure 2.** Percent occurrence of the NRC lexicon emotions in the three parts of *The Alchemist*

*The Alchemist* is an inspirational book that many people turn to during crossroads in their lives. Unsurprisingly, the predominant emotion in all portions of the book is positive. Santiago always carries with him an unwavering trust in his personal destiny. In Part Two, we see more negative emotions and fear due to some of the challenges along the journey. Finally, in the epilogue, we see lots of anticipation as Santiago's journey draws to a conclusion. 


### **Citation:** 
NRC emotion lexicon. Saif Mohammad and Peter Turney: http://saifmohammad.com/WebPages/NRC-Emotion-Lexicon.htm
