---
title: "Palmetto binary logistic regression"
description: |
  Using binary logistic regression to predict palmetto species type.
author:
  - name: Laurel Abowd
date: 2021-03-12
output:
  distill::distill_article:
    self_contained: false
    code_folding: hide
---

```{r setup, include=FALSE, warning = FALSE, error = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(GGally)
library(janitor)
library(broom)
library(kableExtra)
```

## **Introduction**

This analysis uses data on two types of Florida palmetto trees to build a species prediction model using binary logistic regression. The model uses palmetto height, canopy length, canopy width, and number of green leaves as predictors for species. The two species of palmettos in this analysis are *Serenoa repens* and *Sabal etonia*.

## **Read in data**

```{r}
# Read in the data as all columns as characters and convert back to numeric and factor as needed

palmetto <- read_csv(here('data/project-02/palmetto.csv'),
                     col_types = cols(.default = 'c')) %>% 
  mutate(height = as.numeric(height), 
         length = as.numeric(length),
         width = as.numeric(width),
         green_lvs = as.numeric(green_lvs), 
         species = as.factor(species))
```

## **Data visualization**

```{r warning = FALSE, fig.align= "center"}
# Data visualizations to explore differences in height, canopy length, canopy length, canopy width, and green leaves for each of the two species. 
# (1 = Serenoa repens 2 = Sabal etonia)

# Height and green leaves for two species
ggplot(data = palmetto, aes(x = green_lvs, y = height)) +
  geom_point(aes(color = species), alpha = 0.5) +
  scale_color_manual(values = c("darkslategray4", "violetred3"), 
                     labels = c("Serenoa repens", "Sabal etonia")) +
  labs(x = "Count of Green Leaves on Tree",
       y = "Maximum Tree Height (cm)",
       title = "Palmetto Maximum Tree Height (cm) vs. Green Leaf Count") +
  theme_minimal()

```
**Figure 1.** Maximum tree height (cm) vs. the count of green leaves on tree for *Seronoa repens* (blue) and *Sabal etonia* (magenta), two dominant palmetto species in south-central Florida from 1981-2017.  Data source: Abrahamson, W.G. (2019) 

While *Serenoa repens* and *Sabal etonia* have similar maximum tree heights (cm), *Sabal etonia* clearly has fewer counts of green leaves than *Serenoa repens*.

```{r warning = FALSE, fig.align = "center"}
# Canopy height and length for the two species
ggplot(data = palmetto, aes(x = width, y = length)) +
  geom_point(aes(color = species), alpha = 0.5) +
  scale_color_manual(values = c("darkslategray4", "violetred3"),
                     labels = c("Serenoa repens", "Sabal etonia")) +
  labs(x = "Widest Canopy Width (cm)",
       y = "Widest Canopy Length (cm)",
       title = "Palmetto Widest Canopy Width (cm) vs. Widest Canopy Length (cm)") +
  theme_minimal()

```
**Figure 2.** Widest canopy length (cm) vs. widest canopy width (cm) perpendicular to canopy length for *Seronoa repens* (blue) and *Sabal etonia* (magenta), two dominant palmetto species in south-central Florida from 1981-2017.  Data source: Abrahamson, W.G. (2019) 

While *Serenoa repens* and *Sabal etonia* have similar maximum canopy width (cm), *Sabal etonia* appears to have wider canopy length than *Serenoa repens*.

## **Binary logistic regression**

**Table 1.** Binary logistic regression results for *Serenoa repens* ('0' factor level) and *Sabal etonia* ('1' factor level) palmetto trees from 1981-2017. Data source: Abrahamson, W.G. (2019)
```{r }
# Binary logistic regression w/ predictor variables of height, length, width, and green leaves
# levels (0 = "1" (Serenoa repens) 1 = "2" (Sabal etonia))

palmetto_blr <- glm(species~ height + length + width + green_lvs,
                            data = palmetto,
                            family = "binomial" )


# Tidy version
blr_tidy <- tidy(palmetto_blr)

# Finalize table
blr_tidy %>% 
  kable(col.names = c("Variable", "Coefficient", "Standard Error", "Z Statistic", "P Value")) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

## **Model classification**

**Table 2.** Number of times the binomial linear regression model correctly predicts palmetto species of either *Serenoa repens* or *Sabal etonia* based on predictor variables of height, canopy length, canopy width, and number of green leaves. Data source: Abrahamson, W.G. (2019)
```{r warning = FALSE, message = FALSE}
# Percentage of time palmettos would be classified as Sabal etonia based on blr model
blr_fitted <- palmetto_blr %>% 
  broom::augment(type.predict = "response")

# Wrangle blr_fitted so that there is a new column with the predicted palmetto species using a 50% accuracy cutoff. Then create a table showing how many trees were predicted correctly for each species

blr_predict <- blr_fitted %>% 
  filter(.fitted != "NA") %>% 
  mutate(species_prediction = case_when(
    .fitted < .5 ~ "1", 
    .fitted >= .5 ~ "2"
  ))

# Create table
predict_counts <- blr_predict %>%
  mutate(prediction = case_when(
    species == species_prediction ~ "correct",
    species != species_prediction ~ "incorrect"
  )) %>% 
    select(c("species", "species_prediction", "prediction")) %>% 
  group_by(species, prediction) %>% 
  summarize(n = n()) %>% 
  mutate(species = case_when(
    species == "1" ~ "Serenoa repens",
    species == "2" ~ "Sabal etonia"
  )) %>% 
  pivot_wider(names_from = "prediction", values_from = "n") %>% 
  mutate(correctly_predicted = round(((correct)/(correct + incorrect)),digits = 3))

# Finalize table
predict_counts %>% 
  kable(col.names = c("Species", 
                      "Correctly Classified", 
                      "Incorrectly Classified", 
                      "% Correctly Classified")) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "center")

```

## **Summary**

The binary logistic regression model performed very well. Using the test data, the model correctly predicted 90.8% of the *Serenoa repens* palmettos and 92.3% of the *Sabal etonia* palmettos.

### **Citation**
Abrahamson, W.G. 2019. Survival, growth and biomass estimates of two dominant palmetto species of south-central Florida from 1981 - 2017, ongoing at 5-year intervals ver 1. Environmental Data Initiative. https://doi.org/10.6073/pasta/f2f96ec76fbbd4b9db431c79a770c4d5

