---
title: "Bootstrapping"
description: |
  Using bootstrapping to find confidence interval.
author:
  - name: Laurel Abowd
date: 2021-03-14
output:
  distill::distill_article:
    self_contained: false
    code_folding: hide

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(boot)
```

## **Introduction**
This analysis uses bootstrapping to create a 95% confidence interval for the percentage of University of California Santa Barbara (UCSB) community members identifying as nonbinary/genderqueer that have personally experienced, "exclusionary, offensive, hostile or intimidating conduct." Data is from the 2014 UCSB Campus Climate Project survey.

## **Build dataset**
22 of 36 UCSB community members (61%) identifying as nonbinary/genderqueer responded that they had personally experienced, "exclusionary, offensive, hostile or intimidating conduct"

```{r}
# Create vector to relect this data
survey_data <- vector(length = 36)
values <- c(0, 1)

survey_data[1:22] <- unique(values)[2]
survey_data[23:36] <- unique(values)[1]

# 1 = experienced intimidating conduct
# 0 = did not experience intimidating conduct

```

## **Build function**

```{r}
# Function of proportion of sample that experienced intimidating conduct
prop_fun <- function (x,i) {sum(x[i])/length(x[i])}

```

## **Bootstrap 10,000 samples**

```{r}
survey_data_10k <- boot(survey_data,
                        statistic = prop_fun, 
                        R = 10000)

# survey_data_10k

# survey_data_10k$t
  
```

## **Plot data** 

```{r}
# Create dataframe
survey_data_10k_df <- data.frame(bs_prop = survey_data_10k$t)

# Histogram
ggplot(data = survey_data_10k_df, aes(x = bs_prop)) +
  geom_histogram(bins = 22, fill = "#275070") +
  theme_minimal() +
  labs(y = "Count",
       x = "Proportion",
       title = "Bootstraped Survey Data: n = 10,000")
```

## **Confidence interval**
```{r}
boot.ci(survey_data_10k, conf = 0.95, type = "perc")
```

In the 2014 UCSB Campus Climate Project Final Report, 22 of the 36 of surveyed UCSB community members identifying as nonbinary/genderqueer responded that they had personally experienced, "exculsionary, offensive, hostile or intimidating conduct," or a  proportion of 0.61 (n = 36), with a bootstrapped 95% confidence interval of [0.44, 0.78] (n = 10,000 bootstrap samples).

### **Data available at:**
https://campusclimate.ucop.edu/_common/files/pdf-climate/ucsb-full-report.pdf
