---
title: "Cetacean Species Richness in California Bight"
description: |
  Raster analysis
author:
  - name: Laurel Abowd
date: 2021-02-25
output:
  distill::distill_article:
    self_contained: false
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(raster)
library(tidyverse)
library(here)
library(sf)
library(fasterize)
library(maps)
library(mapdata)
library(rnaturalearth)
library(rnaturalearthdata)
library(ghibli)
```

## **Raster Analysis**

```{r}
# Create function to convert cetacean species of 0 or 1 using a threshold of 0.6 for prescence 

species_present <- function(x, thresh = .6) {
 y <- ifelse(x >= thresh, 1, 0)
  return(y)
}

# Read in  the data
cetaceans <- here("data", "project-01", "ca_cetaceans")

cetaceans_files <- dir(cetaceans, full.names = TRUE, pattern = "*.tif")

cetaceans_data <- stack(cetaceans_files) 

# Use function to convert to 0 or 1

cetaceans_present <- calc(cetaceans_data, fun = species_present)

# Find summation of species present by adding up the layers

cetaceans_sum <- calc(cetaceans_present, fun = sum, na.rm = TRUE)

```

## **Finalize Map**

```{r}
# Create finalized map with ggplot

# Convert to data frame
cetaceans_df <- raster::rasterToPoints(cetaceans_sum) %>%
  as.data.frame()

# Get world data
world <- ne_countries(scale = "medium", returnclass = "sf")

# Map with ggplot

ggplot() +
  geom_raster(data = cetaceans_df, aes(x = x, y = y, fill = layer)) +
  scale_fill_ghibli_c("MononokeMedium", direction = +1) +
  geom_sf(data = world) +
  coord_sf(xlim = c(-125, -115), ylim = c(32, 38)) +
  theme_minimal() +
  labs(fill = "Cetacean species\npresent", 
       x = "longitude",
       y = "latitude",
       title = "Cetacean Species Richness in the California Bight")


```

**Figure 1.** Cetacean species richness raster map of the California bight. Color indicates number of species present. A threshold of 0.6 was used to determine the presence or absent of the species. If a species had a greater than 0.6 probability of being presence in that area, it was considered to be present. Data: Kaschner et al. (2016)



### **Data Citation:**   
Kaschner, K., Rius-Barile, J., Kesner-Reyes, K., Garilao, C., Kullander, S., Rees, T., & Froese, R. (2016). AquaMaps: Predicted range maps for aquatic species. www.aquamaps.org

### **Color Palette:**  
Ewen Henderson (2020). ghibli: Studio Ghibli Colour Palettes. R package version 0.3.2.
https://CRAN.R-project.org/package=ghibli 
